### 2015-07-27 doc/thesis/typedlua.tex
Fixing text to mention that tables are tagged and not variables

Typed Lua tags a variable that holds a table type as either
\emph{unique}, \emph{open}, \emph{fixed}, or \emph{closed}.
If a variable gets its type from a table constructor then it is
\emph{unique}, otherwise it is \emph{closed}.
Any alias to \emph{unique} variables makes them \emph{open}, as we use the
tag \emph{open} to keep track of \emph{unique} variables that are aliased.
In the previous example, \texttt{person} has an \emph{unique} type
because it has no alias.
In Section \ref{sec:oop} we will show that we use the tag
\emph{fixed} to describe the type of classes.
Typed Lua also has different subtyping rules that handle these
different tags, which we explain in Section \ref{sec:subtyping}.

Typed Lua uses three rules to decide whether it is sound to
allow field assignment to change the type of \emph{unique} and
\emph{open} variables:
the variable must be local to the current block,
the new type must only add new fields, and
the variable cannot have been assigned in another function.

Before discussing each one of these rules, it is worth discussing
why Typed Lua needs \emph{unique} and \emph{open} table types.
Even though both \emph{unique} and \emph{open} table types
allow the refinement of table types, we need both tags to
increase usability and to avoid some unsafe behaviors.

If we did not have \emph{unique} table types and \emph{open}
table types had the same behavior that they have now,
we would not be able to type check some safe constructions,
like in the following example:
\begin{verbatim}
    local t:{"x":integer, "y":integer?} = { x = 1, y = 2 }
\end{verbatim}

@@

We can create aliases to \emph{open} variables, but they are always \emph{closed}.
Any mutation in the original reference and in the aliased reference
is not a problem, as the type of the original reference can only
add new fields.
For mutable fields this means that the type of a field cannot
change after it is added to the type of a table.
In our previous example, the second line changes the type of
\texttt{person} from \emph{unique} to \emph{open} and makes the type of
\texttt{bogus} \emph{closed}.

>>>>

Typed Lua tags table types as either \emph{unique}, \emph{open},
\emph{fixed}, or \emph{closed} to identify which operations are safe.
The table constructor describes \emph{unique} table types,
as they are tables that we know all the fields and that there is
no alias to it.
Any alias to \emph{unique} tables makes them \emph{open},
as we use this tag to keep track of \emph{unique} aliasing.
In the previous example, \texttt{person} has an \emph{unique} type
because it has no alias.
In Section \ref{sec:oop} we will show that we use the tag
\emph{fixed} to describe the type of classes.
Type annotations define \emph{closed} table types, as this tag
describes table types that do not provide any guarantees about keys
that do not appear in the table type.
Typed Lua also has different subtyping rules that handle these
different tags, which we explain in Section \ref{sec:subtyping}.

Typed Lua uses three rules to decide whether it is sound to
allow field assignment to change the type of \emph{unique} and
\emph{open} tables:
the table must assigned to a local variable to the current block,
the new type must only add new fields, and
the variable cannot have been assigned in another function.

Even though both \emph{unique} and \emph{open} table types
allow the refinement of table types, we need both tags to
increase usability and to avoid some unsafe behaviors.
If we did not have \emph{unique} table types and \emph{open}
table types had the same behavior that they have now,
we would not be able to type check some safe constructions,
like in the following example:
\begin{verbatim}
    local t:{"x":integer, "y":integer?} = { x = 1, y = 2 }
\end{verbatim}

@@

We can create aliases to \emph{open} tables, but they are always \emph{closed}.
Any mutation in the original reference and in the aliased reference
is not a problem, as the type of the original reference can only
add new fields.
For mutable fields this means that the type of a field cannot
change after it is added to the type of a table.
In our previous example, the second line changes the type of
\texttt{person} from \emph{unique} to \emph{open} and makes the type of
\texttt{bogus} \emph{closed}.
